---
title: 'Homework #3'
author: "Ashley Kang"
date: "10/10/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

devtools::install_github("p8105/p8105.datasets")
```

## Problem 1
This problem uses the BRFSS data. First, do some data cleaning:

* Format the data to use appropriate variable names;
* Focus on the “Overall Health” topic
* Include only responses from “Excellent” to “Poor”
* Organize responses as a factor taking levels ordered from “Excellent” to “Poor”

```{r load_clean_data_1}
library(p8105.datasets)

brfss_data = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(state = locationabbr, county = locationdesc) %>% 
  filter(topic == "Overall Health") %>% 
  mutate(response = forcats::fct_relevel(response, c("Excellent", "Very good", "Good", "Fair", "Poor"))) 
```

##### In 2002, which states were observed at 7 locations?
```{r states_location_1}
brfss_data_2002 = brfss_data %>%
  filter(year == "2002") %>%
  group_by(state) %>% 
  summarize(n_locations = n_distinct(county)) %>% 
  filter(n_locations == "7")
```


##### Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.
```{r spaghetti_plot_1}
brfss_data %>% 
  group_by(state, year) %>% 
  summarize(n_locations = n()) %>% 
  ggplot(aes(x = year, y = n_locations, color = state)) +
  labs(
    title = "Spaghetti plot of number of locations in each state from 2002 - 2012",
    x = "Year",
    y = "Number of locations") +
  geom_line() +
  viridis::scale_color_viridis(discrete = TRUE) +
  theme(legend.position = "right")
```

##### Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.
```{r mean_sd_1}
brfss_data %>% 
  filter(year == 2002 | year == 2006 | year == 2010, state == "NY") %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  group_by(state, year) %>% 
  summarize(mean_excellent = mean(excellent, na.rm = TRUE),
            sd_excellent = sd(excellent, na.rm = TRUE)) %>% 
  knitr::kable()
```

##### For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.
```{r distribution_state_avg_1}
brfss_data %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  group_by(state, year) %>% 
  summarize(mean_excellent = mean(excellent, na.rm = TRUE),
            mean_very_good = mean(very_good, na.rm = TRUE),
            mean_good = mean(good, na.rm = TRUE),
            mean_fair = mean(fair, na.rm = TRUE),
            mean_poor = mean(poor, na.rm = TRUE)) %>% 
  gather(key = mean_variable, value = mean_value, mean_excellent:mean_poor) %>% 
  ggplot(aes(x = year, y = mean_value, color = state)) +
  labs(
    title = "Five panel plot showing distribution of state-level averages over time",
    x = "Year",
    y = "Mean Value") +
  geom_line() +
  facet_grid(~mean_variable)  
```

