---
title: 'Homework #3'
author: "Ashley Kang"
date: "10/10/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)


devtools::install_github("p8105/p8105.datasets")
```

## Problem 1
This problem uses the BRFSS data. First, do some data cleaning:

* Format the data to use appropriate variable names;
* Focus on the “Overall Health” topic
* Include only responses from “Excellent” to “Poor”
* Organize responses as a factor taking levels ordered from “Excellent” to “Poor”

```{r load_clean_data_1}
library(p8105.datasets)
data(brfss_smart2010)

brfss_data = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(state = locationabbr, county = locationdesc) %>% 
  filter(topic == "Overall Health") %>% 
  mutate(response = forcats::fct_relevel(response, c("Excellent", "Very good", "Good", "Fair", "Poor"))) 
```

##### In 2002, which states were observed at 7 locations?
```{r states_location_1}
brfss_data_2002 = brfss_data %>%
  filter(year == "2002") %>%
  group_by(state) %>% 
  summarize(number_locations = n_distinct(county)) %>% 
  filter(number_locations == "7")
```

There are 3 states that were observed at 7 locations in 2002: Connecticut (CT), Florida (FL), and North Carolina (NC).

##### Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.
```{r spaghetti_plot_1}
brfss_data %>% 
  group_by(year, state) %>% 
  summarize(number_locations = n()) %>% 
  ggplot(aes(x = year, y = number_locations, color = state)) +
  labs(
    title = "Number of locations in each state, 2002-2012",
    x = "Year",
    y = "Number of locations") +
  geom_line() +
  theme(legend.position = "right") +
  viridis::scale_color_viridis(name = "state", discrete = TRUE)
```

##### Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.
```{r mean_sd_1}
brfss_data %>% 
  filter(year == 2002 | year == 2006 | year == 2010, state == "NY") %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  group_by(state, year) %>% 
  summarize(mean_excellent = mean(excellent, na.rm = TRUE),
            sd_excellent = sd(excellent, na.rm = TRUE)) %>% 
  knitr::kable()
```

##### For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.
```{r distribution_state_avg_1}
brfss_data %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  group_by(state, year) %>% 
  summarize(mean_excellent = mean(excellent, na.rm = TRUE),
            mean_very_good = mean(very_good, na.rm = TRUE),
            mean_good = mean(good, na.rm = TRUE),
            mean_fair = mean(fair, na.rm = TRUE),
            mean_poor = mean(poor, na.rm = TRUE)) %>% 
  gather(key = mean_variable, value = mean_value, mean_excellent:mean_poor) %>% 
  ggplot(aes(x = year, y = mean_value, color = state)) +
  labs(
    title = "Distribution of state-level averages over time per response",
    x = "Year",
    y = "Average proportion") +
  geom_point() +
  geom_line() +
  facet_grid(~mean_variable) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "bottom")
```

## Problem 2
This problem uses the Instacart data.

##### Loading and cleaning Instacart data
```{r loading_cleaning_data_2}
instacart_data = instacart %>% 
  janitor::clean_names()

head(instacart_data)
tail(instacart_data)
```

The dataset `instacart` consists of **`r nrow(instacart_data)` rows** and **`r ncol(instacart_data)`** columns. Each row is a single product from an order and corresponds to information on the `product_name`, `grocery aisle`, `department`, if the product was `reordered`, day of the week and hour the product was ordered (`order_hour_of_day`), and the `user_id` who ordered it. There are **`r nrow(distinct(instacart_data, user_id))` distinct users** who ordered **`r nrow(distinct(instacart_data, product_id))` products** from **`r nrow(distinct(instacart_data, department_id))` distinct departments**.


##### How many aisles are there, and which aisles are the most items ordered from?
```{r aisles_2}
number_aisles = 
  instacart_data %>% 
  distinct(aisle_id) %>% 
  nrow()

instacart_data %>%
  group_by(aisle) %>%
  summarize(number_per_aisle = n()) %>%
  arrange(desc(number_per_aisle)) %>%
  head() %>%
  knitr::kable()
```

There are `r number_aisles` distinct aisles, of which the top 6 aisles with the most items ordered from are:

* Fresh vegetables - 150,609 items ordered
* Fresh fruits - 150,473 items ordered
* Packaged vegetables fruits - 78,493 items ordered
* Yogurt - 55,240 items ordred
* Packaged cheese - 41,699 items ordered
* Water seltzer sparkling water - 36,617 items ordered

#####  Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.
```{r plot_aisles_2}
instacart_data %>%
  group_by(aisle, department) %>%
  summarize(number_per_aisle = n()) %>%
  ungroup() %>%
  mutate(aisle = reorder(aisle, desc(number_per_aisle))) %>% 
  ggplot(aes(x = aisle, y = number_per_aisle, color = department, fill = department)) + 
  labs(
    title = "Number of items ordered per aisle",
    x = "Aisle",
    y = "Number of items") +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        legend.position = "bottom")
  
```

##### Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”
```{r table_most_pop_aisles_2}
instacart_data %>% 
  filter(aisle == "Baking Ingredients" | aisle == "Dog Food Care" | aisle == "Packaged Vegetables Fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(number_per_aisle = n()) %>% 
  top_n(1, number_per_aisle) %>% 
  arrange(desc(number_per_aisle)) %>% 
  knitr::kable()
```

##### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r table_apple_coffee_2}
instacart_data %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_order_hour = mean(order_hour_of_day)) %>% 
  spread(key = order_dow, value = mean_order_hour) %>% 
  setNames(c("Product", "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")) %>%
  knitr::kable()
```

## Problem 3

This problem uses the NY NOAA data.

```{r loading_data_3}
data(ny_noaa)

ny_noaa_data = ny_noaa %>% 
  janitor::clean_names()

head(ny_noaa_data)
tail(ny_noaa_data)

# Missing
missing_precip = ny_noaa_data %>%
  filter(is.na(prcp))

missing_snow = ny_noaa_data %>%
  filter(is.na(snow))

missing_snow_depth = ny_noaa_data %>%
  filter(is.na(snwd))

missing_tmax = ny_noaa_data %>%
  filter(is.na(tmax))

missing_tmin = ny_noaa_data %>%
  filter(is.na(tmin))

# Precipitation
avg_precip = mean(ny_noaa_data$prcp, na.rm = TRUE)

# Snowfall
avg_snow = mean(ny_noaa_data$snow, na.rm = TRUE)

# Snow-depth
avg_snow_depth = mean(ny_noaa$snwd, na.rm = TRUE)
```

The dataset `ny_noaa` contains **`r nrow(ny_noaa_data)` rows** and **`r ncol(ny_noaa_data)` columns**. Each row corresponds to a weather station's report from the corresponding date. The average amount of precipitation is **`r avg_precip` tenths of mm**. The average amount of snowfall is **`r avg_snow` mm**. The average snow depth is **`r avg_snow_depth` mm**. Missing data, particularly for variables `snwd`, `tmax`, and `tmin` is an issue in this dataset, as the proportion of missing `prcp` data is **`r nrow(missing_precip)/nrow(ny_noaa_data)*100`%**, missing `snow` data is **`r nrow(missing_snow)/nrow(ny_noaa_data)*100`%**, missing snow-depth data is **`r nrow(missing_snow_depth)/nrow(ny_noaa_data)*100`%**, missing `tmax` data is **`r nrow(missing_tmax)/nrow(ny_noaa_data)*100`%**, and missing `tmin` data is **`r nrow(missing_tmin)/nrow(ny_noaa_data)*100`%**. 
